\name{}: machine learning over interaction networks and ontology-based background knowledge predicts drug--target interactions

### 2021-04-27 paper/main.tex
.

\begin{figure}[!tpb]
	\centering
	\includegraphics[width=0.9\columnwidth]{figures/model_workflow.png}
	\caption{Full DTI prediction model based on the pretrained
          learnable feature transformations (LFT) for either molecular
          structure or ontology based features. The transformed
          protein representations are added to each corresponded
          protein as node features for the graph convolutional
          steps.}
	\label{fig:ModelWorkflow}
\end{figure}

@@

Considering Macro AUCs, \name{} improves significantly ($p<10^{-4}$,
one-sided t-test) over MolTrans ($t=36.6$), DeepDTI ($t=40.5$),
DeepDTA ($t=36.6$), and DeepConv-DTI ($t=40.5$) on the BioSnap
dataset, and significantly ($p<10^{-4}$, one-sided t-test) over DTINet
($t=20.8$) and DTIGEMS+ ($t=18.0$) on the Yamanishi dataset; DTI-CDF
performs better than \name{} ($p=0.1, t=1.3$, one-sided t-test) \todo{significance}.  We further
perform another one-tailed t-test to compare $MicroAUC_p$
performance. We find that \name{} improved significantly
($p < 10^{-4}$) over MolTrans($t=21.2$), DeepDTI ($t=31.0$), DeepDTA
($t=33.4$), and DeepConv-DTI ($t=23.7$) on the BioSnap dataset, and
over DTINet ($t=29.6$), DTIGEMS+ ($t=28.0$), and DTI-CDF ($t=9.5$) on
the Yamanishi dataset.

In all our experiments, \name{} improves over all other methods with
respect to $MicroAUC_p$, demonstrating that \name{} can identify drugs
that target a specific protein more accurately than other
methods. Several methods achieve a higher MacroAUC than \name{}, in
particular when evaluated using a drug--target pair splitting scheme;
our results with the ``na\"ive'' prediction method show that it may be
possible that models trained on a drug--target split utilize certain
biases in the dataset without necessarily producing novel biological
insights.

@@

\todo[inline]{Why is having only having 20\% of drugs reasonable -> significance test in results}
One major component of our experiments was to determine how the
information that is available to a machine learning model during
training affects the performance of the model. Similarly to previous
work \cite{MolTrans2020, DeepConvDTI2019}, we find significant
differences in predictive performance across different splitting
schemes.

>>>>

\begin{figure}[!tpb]
	\centering
	\includegraphics[width=0.8\columnwidth]{figures/model_workflow.png}
	\caption{Full DTI prediction model based on the pretrained
          learnable feature transformations (LFT) for either molecular
          structure or ontology based features. The transformed
          protein representations are added to each corresponded
          protein as node features for the graph convolutional
          steps.}
	\label{fig:ModelWorkflow}
\end{figure}

@@

Considering Macro AUCs, \name{} improves significantly ($p<10^{-4}$,
one-sided t-test) over MolTrans ($t=36.6$), DeepDTI ($t=40.5$),
DeepDTA ($t=36.6$), and DeepConv-DTI ($t=40.5$) on the BioSnap
dataset, and significantly ($p<10^{-4}$, one-sided t-test) over DTINet
($t=20.8$) and DTIGEMS+ ($t=18.0$) on the Yamanishi dataset; DTI-CDF
performs better than \name{} ($p=0.1, t=1.3$, one-sided t-test).  We further
perform another one-tailed t-test to compare $MicroAUC_p$
performance. We find that \name{} improves significantly
($p < 10^{-4}$) over MolTrans($t=21.2$), DeepDTI ($t=31.0$), DeepDTA
($t=33.4$), and DeepConv-DTI ($t=23.7$) on the BioSnap dataset, and
over DTINet ($t=29.6$), DTIGEMS+ ($t=28.0$), and DTI-CDF ($t=9.5$) on
the Yamanishi dataset.

In all our experiments, \name{} improves over all other methods with
respect to $MicroAUC_p$, demonstrating that \name{} can identify drugs
that target a specific protein more accurately than other
methods. Some methods achieve a higher MacroAUC than \name{}, in
particular when evaluated using a drug--target pair splitting scheme;
our results with the ``na\"ive'' prediction method show that it may be
possible that models trained on a drug--target split utilize certain
biases in the dataset without necessarily producing novel biological
insights.

@@

%\todo[inline]{Why is having only having 20\% of drugs reasonable -> significance test in results}
One major component of our experiments was to determine how the
information that is available to a machine learning model during
training affects the performance of the model. Similarly to previous
work \cite{MolTrans2020, DeepConvDTI2019}, we find significant
differences in predictive performance across different splitting
schemes.
